# Numerical Data {#NUMDATA}


## Objectives

1) Define and use properly in context all new terminology.  
2) Generate in `R` summary statistics for a numeric variable including breaking down by cases.  
3) Generate in `R` appropriate graphical summaries of numerical variables.  
4) Be able to interpret and explain output both graphically and numerically.  



## Homework  

### Problem 1  

**Mammals exploratory**  

Data were collected on 39 species of mammals distributed over 13 orders.  The data is in the `openintro` package as `mammals`

a. Using help, report the units for the variable `BrainWt`.  
```{r eval=FALSE}
?mammals
```


b. Using `inspect` how many variables are numeric?  

```{r warning=FALSE}
inspect(mammals)
```

c. What type of variable is `danger`?

Categorical

d. Create a histogram of `total_sleep` and describe the distribution.  

```{r warning=FALSE}
gf_histogram(~total_sleep,data=mammals,binwidth = 2)
```

```{r warning=FALSE}
gf_dens(~total_sleep,data=mammals)
```

The distribution is unimodal and skewed to the right. It appears it is centered around the value of 11.

e. Create a boxplot of `life_span` and describe the distribution.  

```{r warning=FALSE}
gf_boxplot(~life_span,data=mammals)
```


f. Report the mean and median life span of a mammal.  

```{r}
mean(~life_span,data=mammals,na.rm=TRUE)
```

```{r}
median(~life_span,data=mammals,na.rm=TRUE)
```

g. Calculate the summary statistics for `LifeSpan` broken down by `Danger`.

```{r}
favstats(life_span~danger,data=mammals)
```

### Problem 2  

**Mammals life spans**  

Continue using the `mammals` data set.

a. Create side-by-side boxplots for `life_span` broken down by `exposure`. Note: you will have to change `exposure` to a `factor()`. Report on any findings.  

```{r warning=FALSE}
mammals %>%
gf_boxplot(life_span~factor(exposure))
```

Mammals who are more exposed have a longer life span. There must be a confounding variable, maybe the size of the animal or the `danger` variable.


b. What happened to the median and third quartile in exposure group 4?

```{r}
favstats(life_span~factor(exposure),data=mammals)
```

The median and third quartile are equal in exposure group 4. There are a large number of the observed mammals with the same life span in this group.

c. Create faceted histograms. What are the shortcomings of this plot?

```{r warning=FALSE}
gf_histogram(~life_span,color=~factor(exposure),data=mammals)
```

This is awful.

```{r warning=FALSE}
gf_histogram(~life_span|factor(exposure),data=mammals)
```

Not enough data for each histogram; some of the histograms provide little to no information. Let's do denisty plots.

```{r warning=FALSE}
gf_dens(~life_span,color=~factor(exposure),data=mammals)
```

```{r warning=FALSE}
gf_dens(~life_span|factor(exposure),data=mammals)
```

Which do you think is the best graph?

d. Create a new variable `exposed` that is a factor with level `Low` if exposure is `1` or `2` and `High` otherwise.

```{r}
mammals <- mammals %>%
  mutate(exposed=factor(ifelse((exposure==1)|(exposure==2),"Low","High")))
```


```{r warning=FALSE}
inspect(mammals)
```


e. Repeat part c with the new variable.

```{r warning=FALSE}
gf_dens(~life_span,color=~exposed,data=mammals)
```

### Problem 3  

**Mammals life spans continued**  

a. Create a scatterplot of life span versus length of gestation.  

```{r warning=FALSE}
mammals %>%
gf_point(life_span~gestation)
```


b. What type of an association is apparent between life span and length of gestation?   

It is a weak positive association.

c. What type of an association would you expect to see if the axes of the plot were reversed, i.e. if we plotted length of gestation versus life span?  

The same as this.  Since this is observational data there is no reason to believe there is a causal relationship just by looking at the data. Switching the axis will preserve the association.

d. Create the new scatterplot suggested in c.  

```{r warning=FALSE}
mammals %>%
gf_point(gestation~life_span)
```

e. Are life span and length of gestation independent? Explain your reasoning. 

No there is an association and it appears to be linear. If the plot looked like a "shotgun" blast, we would consider the variables to be independent. However, remember there may be confounding variables that could impact the association between these variables.

