# Multiple Linear Regression {#LRMULTI}

## Objectives

1) Create and interpret a model with multiple predictors and check assumptions.  
2) Generate and interpret confidence intervals for estimates.  
3) Explain adjusted $R^2$ and multi-collinearity.  
4) Interpret regression coefficients for a linear model with multiple predictors.    
5) Build and interpret models with higher order terms.

## Homework 

### Problem 1  

The `mtcars` dataset contains average mileage (mpg) and other information about specific makes and models of cars. (This dataset is built-in to `R`; for more information about this dataset, reference the documentation with `?mtcars`). 

a. Build and interpret the coefficients of a model fitting `mpg` against displacement (`disp`), horsepower (`hp`), rear axle ratio (`drat`), and weight in 1000 lbs (`wt`). 

```{r}
cars_mod<-lm(mpg~disp+hp+drat+wt,data=mtcars)
summary(cars_mod)
```

$$
\E(\text{mpg})=29.15+0.004*\text{disp}-0.035*\text{hp}+1.768*\text{drat}-3.480*\text{wt}
$$

Each coefficient represents the expected increase in `mpg` for a unit increase in the respective variable, leaving all other variables constant. 

b. Given your model, what is the expected, average, mpg for a vehicle with a displacement of 170, a horsepower of 100, a `drat` of 3.80 and a wt of 2,900 lbs. Construct a 95% confidence interval and prediction interval for that expected mpg. 

```{r}
predict(cars_mod,newdata=data.frame(disp=170,hp=100,drat=3.8,wt=2.9),interval="confidence")
predict(cars_mod,newdata=data.frame(disp=170,hp=100,drat=3.8,wt=2.9),interval="prediction")
```

c. Repeat part (b) with a bootstrap for the confidence interval.

```{r cache=TRUE}
set.seed(732)
results <- do(1000)*lm(mpg~disp+hp+drat+wt,data=resample(mtcars))
```

```{r}
head(results)
```


```{r}
results %>%
mutate(pred=Intercept+disp*170+hp*100+drat*3.8+wt*2.9) %>%
cdata(~pred,data=.)
```

### Problem 2 

Is that the best model for predicting mpg? Try a variety of different models. You could explore higher order terms or even interactions. One place to start is by using the `pairs()` function on `mtcars` to plot a large pairwise scatterplot. How high could you get adjusted $R$-squared? Keep in mind that is only one measure of fit. 

Answers will vary, but we tried this and got 0.8694. 

```{r}
summary(lm(mpg~disp+I(disp^2)+hp+I(hp^2)+wt,data=mtcars))
```


